---
title: "          "
author: "Gitanshu Munjal"
date: "March 6th, 2018"
output: 
  html_notebook: 
    highlight: haddock
    theme: spacelab
---
<br><br><br><br>
Clean-up notes can be found [here](http://grmunjal.github.io/presentations/socialcops/Munjal-NRHM-cleanup.nb.html) and Analysis can be found [here](http://grmunjal.github.io/presentations/socialcops/Munjal-NRHM-code.nb.html)
<br><br><hr><br><br>

#### Notes/code for ***Data Analysis***

<li> 
Navigation through the data implemented using 3 core helper [functions](https://github.com/grmunjal/socialcops/tree/master/code/):
<br><br>

1. The ***subject.area()*** function takes a letter input and returns available topics in that part of the survey.

Example:
```{r}
subject.area("A")
```
<br><br>

2. The ***focus.topics()*** function takes a number input and returns available topics in that part of the survey.

Example (say you are interested in topic 1 Ante Natal Care Services ANC from above):

***SCROLL RIGHT ON THE FRAME BELOW***
```{r, fig.width=15}
focus.topics(1)
```
<br><br>

3. The ***focus.data()*** function takes a series of (or single) number input and returns available data for that part of the survey.

Example (say you are interested in vaccination of pregnant women. The above frame tells you that those data start at row 8 and end at row 10 so you can run:

***SCROLL RIGHT ON THE FRAME BELOW***
```{r}
iso.data <- focus.data(8:10)
iso.data
```

<li> These functions together with code that I can use repetitively is then used for the visualizations below. 

<li> The code that appears repeatedly is commented/explained the first time and subsequently just mentioned that it is repeated.

<br><br><br>