---
title: " "
author: "Gitanshu Munjal"
date: "March 6th, 2018"
output: 
  html_notebook: 
    highlight: haddock
    theme: spacelab
---
<br><br><br><br>
Analysis notes can be found [here](http://grmunjal.github.io/presentations/socialcops/Munjal-NRHM-analysis.nb.html) and Analysis can be found [here](http://grmunjal.github.io/presentations/socialcops/Munjal-NRHM-code.nb.html)
<br><br><hr><br><br>


#### Notes/code for ***Data Clean-Up***
<hr>

<li>
Developed a clean [rubric template of the survey](https://github.com/grmunjal/socialcops/tree/master/data)

<li>
Extracted all .xls files given into a [single folder](https://github.com/grmunjal/socialcops/tree/master/data/xls-files/) using bash
```{r}
cd socialcops/data/xls-files

find ../given/ -name '*.xls' -exec cp {} ./ \;

remove=Copy\ of\ 
for filename in  "$remove"*; do mv "$filename" "${filename#"$remove"}"; done
```
<li>
Developed a [Python based parser](https://github.com/grmunjal/socialcops/tree/master/code/) that takes an excel file, fishes out all the data and labels the rows according to the rubric

<li>
Ran parser recursively on all .xls files using bash (less than 1 minute for all files):

```{r}
#Assuming you are in the "xls-files" directory

for filename in *".xls"; do python ../../code/NRHM-parser.py ./"$filename" ../NRHM-MH-tidyrubric.csv; done

mv tidy* ../tidy-files
```



<li>
Output from parser is a "tidy" .csv file that is R readable. File names for the tidy files match the Facility Name inside the spreadsheet. [One file produced per xls provided](https://github.com/grmunjal/socialcops/tree/master/data/tidy-files/). 

<li>
All files merged to form the merged file and column headers added using bash:
```{r}
#Assuming you are in the "tidy-files" directory

cat tidy-* > tidy-NRHM-compiledset.csv

{ echo 'FacilityName,id,SubjectArea,Focus,Description,Grouping,Question,Names,FacilityLevel,Apr.2015,May.2015,Jun.2015,Jul.2015,Aug.2015,Sep.2015,Oct.2015,Nov.2015,Dec.2015,Jan.2016,Feb.2016,Mar.2016'; cat tidy-NRHM-compiledset.csv; } > /tmp/$$ && mv /tmp/$$ tidy-NRHM-compiledset.csv

mv tidy-NRHM-compiledset.csv ../
```


<li> Final file is [here (tidy-NRHM-compiledset.csv)](https://github.com/grmunjal/socialcops/tree/master/data/)
<br><br><br><br>
